---
title: "P8105 HW1 – mk4992"
author: "Maryam Khalid"
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
set.seed(8105)
```

**Problem 1**

```{r libraries}
library(moderndive)  
library(ggplot2)     
```

```{r load-data}
data("early_january_weather")
```

Data Summary 
Rows: `r nrow(early_january_weather)`  
Columns: `r ncol(early_january_weather)`  
Mean temperature: `r round(mean(early_january_weather$temp, na.rm = TRUE), 1)` °F

Key variables (see help in Console with `?early_january_weather`):`time_hour` (date & timestamp), `temp` (°F), `humid` (0–100), `dewp`, `wind_dir`, `wind_speed`, `pressure`, `visib`, etc.

Scatterplot: temp(y) vs time_hour(x); colored by humidity

```{r plot-temp-vs-time, fig.width=8, fig.height=4}
p_temp_time <- ggplot(early_january_weather,
                      aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.7, size = 1.5, na.rm = TRUE) +
  scale_color_continuous(name = "Humidity") +
  labs(
    title = "Temperature over Time (colored by Humidity)",
    x = "Time (hourly)",
    y = "Temperature (°F)"
  ) +
  theme_minimal(base_size = 12)

p_temp_time
```

**Note:** The temperature fluctuates daily from day to night. In addition, when the temperature is higher, the humidity is typically increased as well.

```{r save-plot}
ggsave("temp_vs_time_humid_scatter.png", p_temp_time, width = 8, height = 4, dpi = 300)
```


```{r}
sessionInfo()
```